Fast sparse period estimation

### 2014-06-13 response/response.tex

These theoretical difficulties do not negate the immediate practical applicability of our estimator.  We believe it would be unwise to halt dissemination of these practical results simply because a difficult theoretical problem has not yet been solved.  We agree with the reviewer that the simulations presented are limited. This is entirely due to the 4 page limit required IEEE Signal Processing Letters.  In the paper we use our simulations to motivate the choice $q = 5 f_{\text{max}}$ for the quantisation parameter.  It is natural to question whether this choice also works in other scenarios, for example (as the reviewer suggests), for different noise distributions, and different methods of generating the integers $s_1,\dots,s_N$.  The results of such simulations are displayed in Figures~\ref{plot:poisson} and~\ref{plot:geomuniform} in this document. The simulations are setup as in the paper, but with the following modifications:

\begin{enumerate}
\item In Figure~\ref{plot:poisson} the noise is Gaussian distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right. 
\item In Figure~\ref{plot:geomuniform} the noise if uniformly distributed, $N=200$, and the integers are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$ in the plot on the right. 
\end{enumerate}

Observe that, in both simulations, the choice $q=5 f_{\text{max}}$ is a good one. With this choice the accuracy of the quantised periodogram estimator is similar to the original periodgram estimator (computed without quantisation and FFT).  %That the choice $q=5 f_{\text{max}}$ works in a wide range of scenarios should not come as a

\begin{figure*}[tp] 
  \centering 
  \begin{tikzpicture} 
    \selectcolormodel{gray} 
    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200poisson1}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200poisson1};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5poisson1};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0poisson1};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200poisson1};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200poisson1}; 
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200poisson1};
      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200poisson1};
     \legend{Periodogram theory, LS theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
   \end{axis} 
 \begin{axis}[name=plot2,at={(10.5,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
      \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200poisson10}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200poisson10};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5poisson10};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0poisson10};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200poisson10};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200poisson10}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200poisson10};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200poisson10};
     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
   \end{axis} 
  \end{tikzpicture}  
  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussain noise and $N=200$ observations.  The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically Poisson distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.}\label{plot:poisson}
\end{figure*}

\begin{figure*}[tp]  
  \centering
  \begin{tikzpicture}  
    \selectcolormodel{gray} 
    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=7, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200uniformgeom1}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200uniformgeom1};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5uniformgeom1};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0uniformgeom1};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN200uniformgeom1};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN200uniformgeom1}; 
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200uniformgeom1};
      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200uniformgeom1};
     \legend{Periodogram theory, LS theory, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
   \end{axis}   \begin{axis}[name=plot2,at={(4.7,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200uniformgeom10}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200uniformgeom10};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q1.5uniformgeom10};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN200q5.0uniformgeom10};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN200uniformgeom10};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN200uniformgeom10}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN200uniformgeom10};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN200uniformgeom10};
     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
   \end{axis} 
  \end{tikzpicture}  
  \caption{Mean square period error versus noise variance $\sigma^2$ with uniform noise and $N=200$ observations.  The integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.  The least squares estimator is the most accurate in this simulation. The is predicted by the asymptotic theory~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.}\label{plot:geomuniform}
\end{figure*}

@@

Our paper clearly shows that such limitation do not exist.  The periodgram estimator, computed by quanitisation and FFT and combined with an appropriate optimisation proceedure, is both compuationally efficient and highly accurate.  For moderately large $N$ this is the most computationally efficient proceedure currently available for this problem.  Simultaneously, it is as accurate at the most accurate estimators available, namely the least squares estimator, and the periodgram estimator computed directly without FFT.

In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial. To the authors knowledge, a Cramer-Rao bound is not known in this case.   The \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) supposes a single fixed realisation for $s_1,\dots,s_N$.  This bound is therefore not valid for the simulations we present in Figure~\ref{plot:multipleN} in Section~\ref{sec:simulations} of the paper.

In Figure~\ref{plot:crb} in this document we perform a simulation with integers $s_1,\dots,s_N$. The integers are generated once such $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed.  This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.  Observe that for small noise variance the \emph{clairvoyant} Cramer-Rao bound is similar to the asymptotic variance of the least square and periodogram estimator predicted in~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.  As the noise variance increases the \emph{clairvoyant} Cramer-Rao bound and the asymptotic variances diverge.  As expected, the behaviour of the estimators themselves follows the asymptotic theory, rather than the \emph{clairvoyant} Cramer-Rao bound.

Observe that in the right hand plot the quantised periodogram estimator with coarse quantisation $q = 1.5 f_{\text{max}}$ actually perform well for sufficently small noise variance.  It appears that the fixed integers $s_1,\dots,s_N$ generated in this simulation favour the quantised periodogram estimator.  This is one motivation for performing our simulations with integer $s_1,\dots,s_N$ generated randomly for each Monte-Carlo trial, rather than choosing a single fixed set of integers $s_1,\dots,s_N$.  It avoid the scenario where a particular favourable (or unfavourable) set $s_1,\dots,s_N$ is chosen by chance.

We have not implemented and do not present simulations results with any of the MEA estimators from the papers by Casey and Sadler~\cite{536682,726812}.  The MEA estimators were supersceeded by the Separable Line Search (SLS) estimators of Sidiropoulos, Swami and Sadler~\cite{Sidiropoulos2005} in 2005.  Comparison between the MEA estimator and the SLS are presented in~\cite{Sidiropoulos2005} where the superiority of the SLS estimators is shown.  An enhanced SLS estimator was developed in~\cite{Clarkson2007} and this is the estimator we show simulation results for in our paper and this document.  Even this enhanced SLS estimator performs poorly compared to the least squares and periodogram estimators.  The MEA estimators are no longer state-of-the-art for this problem.

@@

%We prefer to show to plot the MSE of the estimatorsThe least squares estimator is the maximum likelihood estimator, and so, the dashed curve serves a similar type of purpose as the CRB often does in these types of simulations.  Namely, to indicate whether an estimator is performing similarly to what might be considered `optimal', i.e. close the what might theoretically be expected from the maximum likelihood estimator.

\begin{figure*}[tp]  
  \centering
  \begin{tikzpicture}  
    \selectcolormodel{gray}
    \begin{axis}[name=plot1,font=\footnotesize,xmode=log,ymax=4e-1,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.24)}},xlabel style={at={(0.5,0.07)}}, legend style={at={(1.1,1.02)}, anchor=south,legend columns=8, cells={anchor=west,/tikz/every even column/.append style={column sep=0.2cm}}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed1}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed1};
      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed1};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed1};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed1};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/PeriodogramN500fixed1};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.2}] table {code/data/NormalisedSamplingLLSN500fixed1}; 
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed1};
      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed1};
     \legend{Periodogram theory, LS theory, CRB, $q=\tfrac{3}{2}f_{\text{max}}$, $q=5 f_{\text{max}}$, Periodogram, LS, SLS2-new}
     %\node[rotate=17,font=\footnotesize] at (axis cs:6e-4,1e-7) {$N=200$};
   \end{axis}   \begin{axis}[name=plot2,at={(9.5,0)},ymax=4e-1,font=\footnotesize,xmode=log,ymode=log,height=10cm,width=9cm,xlabel={$\sigma^2$},ylabel={MSE},ylabel style={at={(0.06,0.23)}},xlabel style={at={(0.5,0.07)}}]
\addplot[mark=none,color=black] table {code/data/PeriodogramCLTN500fixed10}; 
      \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN500fixed10};
      \addplot[mark=none,color=black,dotted] table {code/data/CRBN500fixed10};
            \addplot[mark=triangle,only marks,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN500q1.5fixed10};
      \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/QuantisedPeriodogramFFTN500q5.0fixed10};
      \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1.1}] table {code/data/PeriodogramN500fixed10};
       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {code/data/NormalisedSamplingLLSN500fixed10}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN500fixed10};
       \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/SLS2newN500fixed10};
     %\node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=200$}; 
   \end{axis} 
  \end{tikzpicture}
  \caption{Mean square period error versus noise variance $\sigma^2$ with Gaussian noise and $N=500$ observations.  A singled fixed set of integers $s_1,\dots,s_N$ are generated so that $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed with mean $\mu = 1$ in the plot on the left and mean $\mu=10$.  This single fixed sequence is used for all of the $500$ Monte-Carlo trials.  In this case the \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) exists and is displayed by the dotted line.  In this simulation the \emph{clairvoyant} Cramer-Rao bound is equal to the variance of the maximum likelihood (i.e. least squares) estimator in the case that the integers $s_1,\dots,s_N$ are known.}\label{plot:crb}
\end{figure*}

### 2014-06-13 response/response.tex

In the simulations we regenerate the integers $s_1,\dots,s_N$ on each Monte-Carlo trial. To the authors knowledge, a Cramer-Rao bound is not known in this case.   The \emph{clairvoyant} Cramer-Rao bound from~\cite{Sidiropoulos2005} (also in \cite{726812,Clarkson2007}) supposes a single fixed realisation for $s_1,\dots,s_N$.  This bound is therefore not valid for the simulations we present in Figure~\ref{plot:multipleN} in Section~\ref{sec:simulations} of the paper.

In Figure~\ref{plot:crb} in this document we perform a simulation with integers $s_1,\dots,s_N$. The integers are generated once such $s_1$ and $s_{n+1} - s_n$ for $n=1,\dots,N-1$ are independent and identically geometrically distributed.  This single fixed sequence is used for all of the $1000$ Monte-Carlo trials.  Observe that for small noise variance the \emph{clairvoyant} Cramer-Rao bound is similar to the asymptotic variance of the least square and periodogram estimator predicted in~\cite{Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}.  As the noise variance increases the \emph{clairvoyant} Cramer-Rao bound and the asymptotic variances diverge.  As expected, the behaviour of the estimators themselves follows the asymptotic theory, rather than the \emph{clairvoyant} Cramer-Rao bound.

Observe that in the right hand plot the quantised periodogram estimator with coarse quantisation $q = 1.5 f_{\text{max}}$ actually perform well for sufficently small noise variance.  It appears that the fixed integers $s_1,\dots,s_N$ generated in this simulation favour the quantised periodogram estimator.  This is one motivation for performing our simulations with integer $s_1,\dots,s_N$ generated randomly for each Monte-Carlo trial, rather than choosing a single fixed set of integers $s_1,\dots,s_N$.  It avoid the scenario where a particular favourable (or unfavourable) set $s_1,\dots,s_N$ is chosen by chance.

We have not implemented and do not present simulations results with any of the MEA estimators from the papers by Casey and Sadler~\cite{536682,726812}.  The MEA estimators were supersceeded by the Separable Line Search (SLS) estimators of Sidiropoulos, Swami and Sadler~\cite{Sidiropoulos2005} in 2005.  Comparison between the MEA estimator and the SLS are presented in~\cite{Sidiropoulos2005} where the superiority of the SLS estimators is shown.  An enhanced SLS estimator (SLS2-new) was developed in~\cite{Clarkson2007} and this is the estimator we show simulations results for.  Even this enhanced SLS estimator performs poorly compared to the least squares and periodogram estimators.  The MEA estimators are no longer state-of-the-art for this problem.

@@

\item\textbf{Comment:}\label{com:rev1:nr}
p. 7, the authors indicate that they have used the Newton-Raphson (NR)
method. More details should be provided here. For instance, what is the exact
NR scheme used in this paper (there are several options)? How have the
authors selected the total number of iterations? Is the quantized periodogram
just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?
Also, do the results displayed in Fig. 2 correspond to the total time
required to compute the estimator (including the NR procedure) or just to the
initialization of the NR procedure?
\\
\textbf{Response:}
We reponse to each of the questions of the reviewer in turn:

\begin{enumerate}
\item \textbf{For instance, what is the exact NR scheme used in this paper (there are several options)?}  The Netwon-Raphson method used proceeds by the iteration
\[
x_{n+1} = x_n - \frac{f^\prime(x_{n})}{f^{\prime\prime}(x_n)}
\]
to find a stationary point of the function $f$ where $f^\prime$ and $f^{\prime\prime}$ are the first and second derivatives of $f$.  To avoid the potential ambguity regarding the definition of the Newton-Raphson method we have, in the new manuscipt, replaced the Newton-Raphson method with the optimisation proceedure due to Brent~\cite[Ch.~5]{Brent_opt_no_derivs_1973}.  This optimiation method is likely to be more familiar with readers because it is used internally by the \texttt{fminbnd} function in the popular MATLAB environment.

BLERG add some text to the paper.

\item \text{How have the authors selected the total number of iterations?} In our implementation iterations are performed until the absolute difference between consequentive iterations is less than $10^{-7}$.

BLERG: Something like the following should be added:

``Brent's method is iterative.  During the method a number of evaluations of the periodogram $I_y$ are made.  The number of calls depends on the grid width $D$, but not on $N$.  Each call requires $O(N)$ operations, and so, the total number of operations required by Brent's method is $O(N)$. In practice we find that that computational cost of Brent's method is far smaller than the cost of either that FFT or chirp-z transform.''

We are only aware of a single   As was stated Section~\zref{sec:simulations}: ``Both the periodogram and quanitised periodogram estimators use the Newton-Raphson method.''  This was the case both for the simulations presented in Figure~\zref{plot:multipleN} and for the computational benchmarks presented in Figure~\zref{plot:benchmark}.

@@

BLERG: In Section~\zref{sec:peri-estim} the text has been modifed to read:

``An iterative procedure, such as the Newton-Raphson method, or Brent's method~\cite{Brent_opt_no_derivs_1973} 
can then be used to compute $\hat{f}$ starting from $\widetilde{f}$.''

\item\textbf{Comment:}
p. 5, in the definition of $v_m$, $b_{m+\ell_1}$ should be replaced by $b_{m+\ell_{\text{min}}}$.
\\
\textbf{Response:}
Thankyou for spotting this error.  A similar error was fixed in equation for $I_z$ directly above this.

\item\textbf{Comment:}
p. 6, ``each of length $L+K-1$'' might be replaced by ``each of length $M = L + K - 1$"
\\
\textbf{Response:}
Respectfully, we have not done this.  The parameter $M$ is reserved for the length of the FFT used for the single Fourier transform method.  We feel it is confusing if $M$ is also used for the length of the transforms required by the chirp-z transform.

@@

\item\textbf{Comment:}
p. 8, $n = 2,\dots,N$ should be $n = 1, \dots, N - 1$
\\
\textbf{Response:}
This error has be fixed.  The same error in the caption of Figure~\zref{plot:multipleN} has been fixed.

\item\textbf{Comment:}
p. 8, it is not completely clear to identify the curves associated with $N = 30$ and $N = 1200$. Can the authors display 4 curves associated with $\mu= 1$, $\mu = 10$, $N = 30$ and $N = 1200$?
\\
\textbf{Response:}
We have connected the data points for the $N=1200$ case by gray lines.  We believe this makes the plot clear.  We are happy to make further adjustments to this plot should the reviewer require them.

\item\textbf{Comment:}
Typos: p. 2, ``anomolies'' should be ``anomalies'', p. 7, ``quanitised'' should be
``quantized''.
\\
\textbf{Response:}
Done.

### 2014-06-13 response/response.tex

\item\textbf{Comment:}\label{com:rev1:nr}
p. 7, the authors indicate that they have used the Newton-Raphson (NR)
method. More details should be provided here. For instance, what is the exact
NR scheme used in this paper (there are several options)? How have the
authors selected the total number of iterations? Is the quantized periodogram
just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?
Also, do the results displayed in Fig. 2 correspond to the total time
required to compute the estimator (including the NR procedure) or just to the
initialization of the NR procedure?
\\
\textbf{Response:}
We reponse to each of the questions of the reviewer in turn:

\begin{enumerate}
\item \textbf{For instance, what is the exact NR scheme used in this paper (there are several options)?}  The Netwon-Raphson method that was used proceeds by the iteration
\[
x_{n+1} = x_n - \frac{f^\prime(x_{n})}{f^{\prime\prime}(x_n)}
\]
to find a stationary point of the function $f$ where $f^\prime$ and $f^{\prime\prime}$ are the first and second derivatives of $f$.  To avoid the potential ambguity regarding the definition of the Newton-Raphson method we have, in the new manuscipt, replaced the Newton-Raphson method with the optimisation proceedure due to Brent~\cite[Ch.~5]{Brent_opt_no_derivs_1973}.  This optimiation method is likely to be more familiar with readers because it is used internally by the \texttt{fminbnd} function in the popular MATLAB environment.

@@

BLERG add some text to the paper.

\item \textbf{How have the authors selected the total number of iterations?} In our implementation iterations are performed until the absolute difference between consequentive iterations is less than $10^{-7}$.

@@

BLERG: Something like the following should be added:

``Brent's method is iterative.  During the method a number of evaluations of the periodogram $I_y$ are made.  The number of calls depends on the grid width $D$, but not on $N$.  Each call requires $O(N)$ operations, and so, the total number of operations required by Brent's method is $O(N)$. In practice we find that that computational cost of Brent's method is far smaller than the cost of either that FFT or chirp-z transform.''

\item \textbf{Is the quantized periodogram
just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?}

\item \textbf{Also, do the results displayed in Fig. 2 correspond to the total time
required to compute the estimator (including the NR procedure) or just to the
initialization of the NR procedure?}

@@

\end{enumerate}

\section*{Reviewer 2}

\begin{enumerate}

\item\textbf{Comment:}
 The identifiability claim at the beginning of sec. II is not obvious and a reference is mandatory. Indeed, referring to eq. (2), let us assume that $s_n = 2 i_n$ where $i_n$ are all non-negative integer. In this case the point process is periodic with period $2 T_0$. This referee does not believe that maximizing the periodogram over $(f_{\text{min}},f_{\text{max}})$ will identify the true frequency (or period). LS methods might have the capabilities to identify the period, but, in any case the claim of idenitifiability needs to be further supported.
\\
\textbf{Response:} The same identifiability requirements are commonly imposed in the literature, see for example, Sidiropoulos~\cite[eq. (2)]{Sidiropoulos2005}.  Suppose that $T_{\text{max}} \geq 2T_{\text{min}}$ and that the true period $T_0$ happened to be contained in the nonempty interval $(2T_{\text{min}}, T_0)$.  Put $T^\prime = T_0/2 \in (T_{\text{min}}, T_{\text{max}})$ and observe that
\[
y_n = T_0 s_n + \theta_0 + w_n = 2 T^\prime s_n + \theta_0 + w_n = T^\prime r_n + \theta_0 + w_n 
\]
where $r_n = 2 s_n \in \ints$.  Since the integers are not known both $T_0$ and $T^\prime$ would describe the observations equally well.  This problem is averted only if we suppose that $T_0 \in (T_{\text{min}}, T_{\text{max}})$ and that  $T_{\text{max}} < 2T_{\text{min}}$.

%Perhaps it is possible to increase the range of identifiability if further assumptions are made regarding the integers $s_1,\dots,s_N$.  We don't consider such propsitions in this paper.

% To see this another way, define least squares objective function $LS : \reals \times \reals \times \ints^N \mapsto \reals$ as
% \[
% LS(T, \theta, z_1,\dots,z_N) = \sum_{n=1}^N (y_n - T z_n - \theta)^2.
% \]
% In the case that the noise variables $w_1,\dots,w_n$ are i.i.d. and normally distributed the minimiser of this objective function with respect to $T$ is the maximum likelihood estimator.  That is,
% \[
% \hat{T} = \arg\min_{T \in (T_{\text{min}}, T_{\text{max}})} \min_{\theta \in \reals} \min_{z_1,\dots,z_N \in \ints} LS(T, \theta, z_1,\dots,z_N)
% \]
% is the maximum likelihood period estimator.

In the paper we have modifed the s BLERG

@@

\item\textbf{Comment:}
Is the equality in the second (un-numbered) equation of page 2 correct?
\\
\textbf{Response:}
Yes.  The terms inside the sum on the left and right are conjugates of one another.  That is,
\[
\sum_{n=1}^N e^{ 2\pi j f y_n} = \sum_{n=1}^N (e^{-2\pi j f y_n})^* = \left( \sum_{n=1}^N e^{ -2\pi j f y_n} \right)^*
\]
where $^*$ denotes the complex conjugate.  So,
\[
I_y(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f y_n} }^2 = \babs{ \sum_{n=1}^N e^{ -2\pi j f y_n} }^2.
\]

\item\textbf{Comment:}
Newton-Rapson iteration to further improve the frequency estimation is advocated in Sec. II. Can a reference supporting this statement be added?
\\
\textbf{Response:}
Please see our response to comment~\ref{com:rev1:nr} of reviewer 1.

### 2014-06-14 response/response.tex

\item\textbf{Comment:}\label{com:rev1:nr}
p. 7, the authors indicate that they have used the Newton-Raphson (NR)
method. More details should be provided here. For instance, what is the exact
NR scheme used in this paper (there are several options)? How have the
authors selected the total number of iterations? Is the quantized periodogram
just used to initialize a Newton-Raphson procedure? Do the MSEs displayed
in Fig. 3 correspond to the estimator obtained after applying the NR algorithm?
Also, do the results displayed in Fig. 2 correspond to the total time
required to compute the estimator (including the NR procedure) or just to the
initialization of the NR procedure?
\\
\textbf{Response:}
We repond to each of the questions of the reviewer in turn:

@@

\item\textbf{Comment:}
p. 8, $n = 2,\dots,N$ should be $n = 1, \dots, N - 1$
\\
\textbf{Response:}
This error has be fixed.  The same error in the caption of Figure~\zref{plot:multipleN} has been fixed.

\item\textbf{Comment:}
p. 8, it is not completely clear to identify the curves associated with $N = 30$ and $N = 1200$. Can the authors display 4 curves associated with $\mu= 1$, $\mu = 10$, $N = 30$ and $N = 1200$?
\\
\textbf{Response:}
We have connected the data points for the $N=1200$ case by gray lines.  We believe this makes the plot clear.  We are happy to make further adjustments to this plot should the reviewer require them.

\item\textbf{Comment:}
Typos: p. 2, ``anomolies'' should be ``anomalies'', p. 7, ``quanitised'' should be
``quantized''.
\\
\textbf{Response:}
Done.

\item\textbf{Comment:} 
Is it necessary to include 11 self citations in this letter? Removing some of
these references would allow to save some space, which could be used to better
explain some points mentioned above.
\\
\textbf{Response:}
IEEE Signal Processing Letters allows a 5th page that can contain only references.  Therefore, we cannot save space by removing references.   We have removed reference [29] from the original manuscript because the material in that paper is also covered in [27].  We have also removed references [33] and [34] from the original manuscript. We feel that the all 35 references in the new version of the paper are highly relevant. The paper would be weakened without them.

@@

%Perhaps it is possible to increase the range of identifiability if further assumptions are made regarding the integers $s_1,\dots,s_N$.  We don't consider such propsitions in this paper.

% To see this another way, define least squares objective function $LS : \reals \times \reals \times \ints^N \mapsto \reals$ as
% \[
% LS(T, \theta, z_1,\dots,z_N) = \sum_{n=1}^N (y_n - T z_n - \theta)^2.
% \]
% In the case that the noise variables $w_1,\dots,w_n$ are i.i.d. and normally distributed the minimiser of this objective function with respect to $T$ is the maximum likelihood estimator.  That is,
% \[
% \hat{T} = \arg\min_{T \in (T_{\text{min}}, T_{\text{max}})} \min_{\theta \in \reals} \min_{z_1,\dots,z_N \in \ints} LS(T, \theta, z_1,\dots,z_N)
% \]
% is the maximum likelihood period estimator.

We have included reference~\cite{Sidiropoulos2005} and simplified the text at the start of Section~\zref{sec:peri-estim} to read:

``To ensure identifiability it is necessary to assume, as in~\cite{Sidiropoulos2005}, that $T_0$ lies in some interval $(T_{\text{min}}, T_{\text{max}})$ where $T_{\text{max}} < 2T_{\text{min}}$.''
